Automated optimization of a course of actions in a stochastic environment

### 2016-05-28 paper.Rmd
te

#General principle

\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
\begin{algorithm*}
\caption{General version}
\begin{algorithmic}
\STATE$\textbf{initialize:}$
\IND$f() \text{: regressor function}$
\IND$x_t \text{: current stochastic environment}$
\IND$\Omega_t(x_t) \text{: set of possible actions}$
\IND$s(x_t) \text{: signal activated or not by the stochastic environment, initialized to 0}$
\IND$W^*=W_0\text{: random weights for the regressor}$
\IND$X\text{: empty training feature matrix}$
\IND$y\text{: empty training target vector}$
\IND$A\text{: empty feature matrix for simulated actions}$
\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
\IND$\Delta = 0 \text{: number of actions}$
\STATE$\textbf{while } \text{the learning process is on:}$
\IND$\text{record }x_t$
\IND$A \text{ = empty matrix}$
\IND$\textbf{for } a' \text{ simulated action in } \Omega_t(x_t)$
\IND[2]$A.append(a')$
\IND$a^*=\arg\max_{a' \in A} f(W^*,A)$
\IND$X.append(a^*)$
\IND$\textbf{if } s(a^*) \text{ != } 0 \text{ :}$
\IND[2]$\textbf{for } a \text{ in } \Delta \text{:}$
\IND[3]$value = \gamma^{a-1}s(a^*)$
\IND[3]$y.append(value)$
\IND[3]$W^* := \arg\min_{W}(l(f(W,X),y))$
\IND[3]$\Delta = 0$
\IND$\textbf{else} \text{:}$
\IND[2]$\Delta \text{ += } 1$
\IND[2]$\text{continue}$
\end{algorithmic}
\end{algorithm*}

The algorithm assumes two underlying function:

@@

- The signal function assumes the existence of one or multiple sensors, or the manual assignement of a value. It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.

>>>>

#General principle
Explain how it works in sentences

\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
\begin{algorithm*}
\caption{General version}
\begin{algorithmic}
\STATE$\text{Initialized:}$
\IND$X\text{: empty }m\text{-columns feature matrix}$
\IND$y\text{: empty target vector}$
\IND$X'\text{: empty }m\text{-columns simulated states matrix}$
\IND$w^*=w_0\text{: random weights for the regressor}$
\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
\IND$\Delta = 0 \text{: number of actions}$
\STATE$\text{Other variables:}$
\IND$f() \text{: regressor function}$
\IND$m \text{: number of features}$
\IND$p_t(x^*) \text{: stochastic process, function of the last existing chosen state}$
\IND$x_t(p_t) \in {\rm I\!R^m} \text{: current state vector}$
\IND$\Omega_t(x_t) \text{: set of possible actions}$
\IND$s(x_t) \in {\rm I\!R} \text{: signal value triggered by the modified state}$
\STATE$\textbf{WHILE } \text{the learning process is on:}$
\IND$\text{record }x_t$
\IND$X' := \text{ empty matrix}$
\IND$\textbf{FOR } x' \text{ in } \Omega_t(x_t)\text{:}$
\IND[2]$\text{append } x' \text{ to } X'$
\IND$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
\IND$\text{append } x^* \text{ to } X$
\IND$\textbf{IF } s(x^*) \text{ != } 0 \text{ :}$
\IND[2]$\textbf{FOR } \delta \text{ in } 1\text{:}\Delta \text{:}$
\IND[3]$value = \gamma^{\delta-1}s(x^*)$
\IND[3]$\text{append } value \text{ to } y$
\IND[2]$w^* := \arg\min_{w}(l(f(w,X),y))$
\IND[2]$\Delta := 0$
\IND$\textbf{ELSE} \text{:}$
\IND[2]$\Delta \text{ += } 1$
\end{algorithmic}
\end{algorithm*}

Explanation of some elements:

- $\Omega$ can be set deterministically. In the context of a game, there is a well determined space of actions. It also can be infinite. A mobile robot can explore the real 3D world in almost infinite ways.

- The signal value function assumes the existence of one or multiple sensors, or the manual assignement of a value. It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.

@@
